Appendix 3: Estimating species interactions

### 2015-11-02 Harris_2015/manuscript-materials/draft.Rmd
fix appendices, I think

\chead{\ifthenelse{\value{page}=1}{}{\textit{Species interactions in Markov networks}}}
\renewcommand{\headrulewidth}{0pt}

@@

I then generated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). As described in Appendix [[X]], I randomly drew
the “true” coefficient values for each replicate landscape
so that most species pairs interacted negligibly, a few pairs interacted very
strongly, and competition was three times more common than facilitation.

The first set of 20-species landscapes, like the landscapes with three species,
were generated directly from a Markov network to ensure that the model could
recover the parameters used to generate the "observed" co-occurrence data.
Then, I added two environmental factors that varied from location to location
across the simulated landscapes, and simulated a new set of co-occurrence data
so that species' $\alpha$ coefficients depended on the local environment.
The latter set of simulated landscapes provide an important test of the methods’
ability to distinguish co-occurrence patterns that were generated from pairwise
biotic interactions from those that were generated by external forces like
abiotic environmental filtering. This task was made especially difficult
because---as with most analyses of presence-absence data for co-occurrence
patterns---the inference procedure did not have access to any information about
the environmental or spatial variables that helped shape the landscape [cf
@connor_checkered_2013].  I generated the final set of landscapes by simulating
the population dynamics of 20 interacting species, as described in Appendix [[X]].

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models,  as described above. For the analyses with 20 species, I added a weak
regularizer (equivalent to a logistic prior with scale 2) to ensure that the
model estimates were always finite [[Appendix]].

>>>>

\chead{\ifthenelse{\value{page}=1}{}{\textit{Inferring species interactions}}}
\renewcommand{\headrulewidth}{0pt}

@@

I then generated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). As described in Appendix 2, I randomly drew
the “true” coefficient values for each replicate landscape
so that most species pairs interacted negligibly, a few pairs interacted very
strongly, and competition was three times more common than facilitation.

The first set of 20-species landscapes, like the landscapes with three species,
were generated directly from a Markov network to ensure that the model could
recover the parameters used to generate the "observed" co-occurrence data.
Then, I added two environmental factors that varied from location to location
across the simulated landscapes, and simulated a new set of co-occurrence data
so that species' $\alpha$ coefficients depended on the local environment.
The latter set of simulated landscapes provide an important test of the methods’
ability to distinguish co-occurrence patterns that were generated from pairwise
biotic interactions from those that were generated by external forces like
abiotic environmental filtering. This task was made especially difficult
because---as with most analyses of presence-absence data for co-occurrence
patterns---the inference procedure did not have access to any information about
the environmental or spatial variables that helped shape the landscape [cf
@connor_checkered_2013].  I generated the final set of landscapes by simulating
the population dynamics of 20 interacting species, as described in Appendix 3.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 4).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models,  as described above. For the analyses with 20 species, I added a weak
regularizer (equivalent to a logistic prior with scale 2) to ensure that the
model estimates were always finite.
